
%% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
%
%\terms{Theory}
%
%\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

Chem-infomatics is largely concerned with the task of mining large chemical data-sets. Typical applications of this include drug discovery, catalyst analysis and experiment predictions. Usually chemicals are represented as "Chemical Fingerprints". Chemical fingerprints are used for identification of chemicals.  The direct comparison of chemical molecules (generally in the form of graph data) is difficult but conversion of the molecules into bit-strings or into vector format (known as the fingerprint of the chemical compound) makes it easier for researchers to do a comprehensive comparison. A primary reason for using chemical fingerprints is scalability. We know that sub-graph isomorphism is an NP-complete problem making it difficult to find similarity between two molecules represented in a graphical format. When the data is represented in a vector string, techniques are more scalable in terms of size of query molecules as well as the time required to accomplish such a similarity detection task.\\

Chemical fingerprints are built based on compound properties like substructure features. In binary fingerprints each bit represents the presence or absence of a specific chemical property. For example the bits could represent the count of individual chemical atoms like carbon, oxygen, hydrogen or the number of saturated or unsaturated aromatic or non-aromatic rings. The reason why substructure presence is an important feature is because it can be used to reduce or filter out candidate mismatches in the sub-graph isomorphism test process. Given two bit-strings or vectors for binary chemical  fingerprints we can intuitively say that the similarity of the two fingerprints is directly proportional to the number of shared bits. \\

These fingerprints/feature vectors are designed by domain experts and typical features include information like number of occurrences of particular substructures, presence of molecules, bonds between the molecules, etc. Generally the data-set is high dimensional and also highly sparse which inherently makes searching and indexing such chemical data sets challenging. To give an example, one of the data-sets we are working on is the PubChem data-set which contains about 264016 number of data points and 785985 total number of unique features. These numbers are representative of the high dimensional nature of the data-set. The sparsity of the data-set is indicated by the fact that the average number of features contained by a data point of the PubChem database is 270.602966.
	
For performing query operation on high dimensional data, it is generally observed that many conventional indexing techniques perform poorly, as compared to linear scan. This is mainly due to the curse of dimensionality . As dimensions increases the data points are scattered further away in space. In some techniques which use pivots to form clusters like the one we will use, it is difficult to form compact groups with a low cluster radius. Since we are unable to form compact representative groups well, more than often a range query requires a search through the entire database of chemical molecules with no chance of pruning occurring.  The results of our range search techniques will be verified by comparing the range search result set against that obtained by a linear scan approach.\\

Similarity between chemical molecules plays an important role in various aspects of biology and chemistry like protein ligand docking, biological activity prediction, reaction site modeling and drug discovery. Furthermore, chemical or molecular similarity  plays a pivotal role in modern techniques used to predict chemical compound properties, designing tailor-made chemicals with a predefined and desirable set of properties and most importantly in managing drug design studies by using large databases containing structures of available chemicals.\\

The process of calculating similarity between chemical compounds involves a number of steps. Firstly each chemical compound which is represented in a graph manner, with neighboring atoms, bond-linkages, substructures, etc have to be represented in a manner where finding similarity becomes simpler. When we represent items in a vector format, our job becomes easier with the use of measures like Jaccard Similarity as well as the Cosine Similarity. Hence the need to convert compounds to a format we described earlier known as fingerprints. \\

The conversion of the compounds to the vector format is known as fingerprinting. In the field of computer science, fingerprinting is a technique that hashes or maps every large data item to a much shorter string, generally a bit vector, also known as its fingerprint. This fingerprint is used to uniquely identify the data item for all purposes, very similar to how fingerprints of humans can be mapped uniquely for every single individual. Fingerprinting as mentioned earlier is used for avoiding transfer and comparison of large bulky data. \\

Various scoring schemes like Tversky, Pearson, Dice, and Kulczynski are available for comparing two given fingerprints \cite{willett2006similarity,swamidass2007bounds}. The most widely used and accepted similarity measure for binary fingerprints is the Tanimoto similarity which is equivalent to the Jaccard Similarity index defined as the size of the intersection of the sets divided by the size of union of the sets. Here the similarity measure would correspond to the number of bits which are 1 for both the sets of the fingerprints divided by the number of bits for which atleast one of the fingerprints is 1. In mathematical terms,for binary fingerprints X and Y ,where $X_i$ is the $i^th$ bit of X we define Tanimoto similarity as	
\begin{equation}
 T_s(X,Y) = \frac{\sum \limits_i X_i \wedge Y_i}
{\sum \limits_i X_i \vee Y_i} 
\end{equation}	

Here we can show that $1-T_s$ is a proper distance metric preserving triangle inequality which will allow us to use index structures which exploit the property, like M-trees. One challenge which we face here is the extension of the said similarity measure /distance metric to non binary fingerprints. Non binary fingerprints are basically feature vectors where each feature value is a count and not necessary signifying absence or presence using bits. \\

%
%MOL2 is a primary molecule format used for the representation of chemical compounds. The MOL2 file displays information specifically about the following, Atom, Molecule, Bond, Substructure and Set. The atom record provides information about the atom names, types, coordinates, and partial charges. The molecule entry recognizes information about the molecule name and number of atoms, bonds, substructures and sets. Similarly the bond information displays the atom identifiers for every bond and the substructure record identifies types of significant substructures seen.\\

The first technique we used is that of M-tree \cite{ciaccia1997indexing} based indexing. In this technique we have proposed a novel pivot selection technique, which results in a efficient search time for range queries. We use triangle inequality bounds on the distance measure to effectively prune out compounds or alternatively include a whole set of compounds in the range search query answer set. We have proposed an indexing technique which runs in linear time against the size of the data-set. We have also proposed methods to take care of out-liers in the data-set during the indexing procedure .\\

We follow this up by embedding the data in to Euclidean space using Lipschitz embedding. We then perform a detailed analysis of the data, and we show that this analysis naturally leads us to inverted indexing. The inverted indexing idea comes from tf-idf (term frequency - inverse document frequency) used in information retrieval and text mining. We propose a unique pruning method which helps in cutting down the candidate list for our result set.  The large size of the data and the sparsity of the data are the primary challenges which we have addressed in this work.\\

To give you an overview of the chapters; In \autoref{RW}, previous work in this field of research have been summarized. We have also outlined the shortcomings of the state of the art techniques which we will address in our work. In \autoref{PF}, we give the formal definition of our problem, study the utilization of the Tanimoto similarity measure and describe the exact threshold similarity search problem which we intend to tackle.

In \autoref{MT}, a detailed description of the M-tree data structure has been given along with the steps proposed for our indexing procedure. We also explore improvisations on the baseline indexing technique and how Lipschitz embedding can be utilized in our indexing method. In \autoref{II}, the use of the inverted index has been analyzed, along with the derivation of the bounds required for the pruning of the features and few other extensions to the baseline indexing technique.

	In \autoref{DA}, we do a comprehensive analysis of the data-sets to help us understand and exploit the nature of the fingerprint data to our advantage. Finally in \autoref{EX}, we discuss the experimental analysis, results and observations of the comparison of our techniques to the state of the art technique on various data-sets. In \autoref{CO}, we present a short summary of our entire work and entail directions for future extensions to our work.



	